---
title: "How boomer works"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{How boomer works}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  error = TRUE # TODO: fix failing chunks
)
```

This vignette summarizes how the internals of boomer work.

## Overview

An important principle of {boomer} is that we don't modify the body of the function
we rig. 

```{r}
rigged_file_ext <- boomer::rig(tools::file_ext)
tools::file_ext
rigged_file_ext
```

Instead we copy the original function but give it a new environment. 
This new environment is a child of the original environment, and is populated with
shims of the functions called by the original function.
We call this environment the mask.

```{r}
# the original environment
environment(tools::file_ext)

# our new environment
env <- environment(rigged_file_ext)
env

# its parent
parent.env(env)

# its content
ls(env)
```

`rig_impl()` does this job and is the main function of the package.
It calls `wrap()`, the other important function, whose mission is to build verbose shims of functions
used in the rigged function. Both are detailed thereafter.

## Individual functions

### `boom()` and `rig()`

`boom()` is a wrapper around `rig_impl()` , it rigs the calling function and runs the call,
it also has some hacky code so we can pipe to `boom()` with {magrittr} (The hack is
not needed for the base pipe)

### `rig_impl()`

Here's the diagram of dependencies of `rig_impl()`

```{r}
flow::flow_view_deps(boomer:::rig_impl, show_imports = "packages")
```

`rig_impl()` :

* Creates a new environment (the mask) as a child of the original environment
* Populates it with shims of `::` and `:::` so when we find `pkg::fun` or `pkg:::fun`
 in the original function we don't wrap their output and not the `::` or `:::` operator
* Populates it with shims of `<-` and `=` so functions created in the original function,
and thus impossible to shim at "rig time", can be made verbose too.
* Populates it with shims of all other called functions, using `wrap()`

### `wrap()`

`wrap()` builds verbose wrapper functions.

Its main aim is to print information directly related to the wrapped function 
(e.g. argument values and execution time).

However it does a couple more things:

* If a wrapper function is the first to be called within a rigged function's
execution, it prints information signaling that we are entering the rigged function,
and sets up an exit hook so we print information that we are exiting it. It detects
this "first call" by the absence of a `..EVALED_ARGS..` variable in the rigged
function's execution environment, which it creates on that occasion.
* It checks if the rigged function's arguments have been evaluated (remember in R
arguments are not evaluated unless their values are requested down the line) and
prints them when they are. This uses that same `..EVALED_ARGS..` variable, a named
logical vector that keeps track of which arguments have been evaled.

### `rig_in_place()`

`rig_in_place()` calls `rig_impl()` on its inputs but unlike `rig()` we want
the rigged functions to be bound in the namespace instead of the original functions.

To do this we unlock the namespace and assign the rigged functions there. We also
replace the other copies a caller might reach: the one in the attached package
environment, the one in the S3 methods table (so S3 dispatch is verbose too), and
the ones imported by other already loaded packages.

Since `rig_in_place()` accepts several functions as arguments, and that they might
call each other, we also make sure we include wrapped versions of our rigged functions
in all the masks.
